\section*{1. Introduction}

We propose a novel cosmological framework in which the universe evolves through discrete recursive cycles, preserving memory via quantum interference across entangled geometries. This model integrates core principles from loop quantum cosmology~\cite{ashtekar2006quantum}, ER=EPR dualities~\cite{maldacena2013cool}, and non-Markovian decoherence~\cite{breuer2002theory}, but introduces a key conceptual advance: the transition between universal cycles is governed by an interference operation across entire spacetime states.

At the heart of the framework is a generalized state vector \( \phi \), encoding spatial dimensions (L, W, H), time (T), color-space frequency modes (R, G, B), amplitude (A), quantum interference (Q), recursive index (I), and observer state (O). The evolution from one universal cycle to the next is described not by linear propagation, but by a logical interference pattern expressed as:

\[
\phi_{n+1} = \phi_{U_1} \oplus \phi_{U_2}
\]

where \( \oplus \) denotes a generalized interference projection analogous to a bitwise XOR. This operation fuses two prior universe-states \( \phi_{U_1} \), \( \phi_{U_2} \) into a new configuration, constrained by entropic consistency and memory fidelity across Einstein-Rosen bridges.

This recursive transition defines a variational action \( \mathcal{A}_n[\phi] \) that balances forward entropy increase with backward coherence retention. The full recursive path integral includes a kernel \( K(\phi, \phi') \) derived from spinfoam amplitude models, modulated by memory-fidelity weights tied to observational constraints such as CMB suppression and quantum gravitational wave signatures.

In this view, the universe becomes a learning system, adjusting its geometry to maximize coherence across cycles. The model provides a falsifiable framework that predicts observable deviations from standard inflationary cosmology — including tensor suppression at low-\( \ell \), gravitational memory echoes, and a bounded entropy growth curve tied to recursive fidelity.

