\section*{9. Theoretical Comparisons and Open Questions}

This framework stands as a synthesis—integrating geometry, quantum information, and recursive dynamics. It builds on Loop Quantum Cosmology~\cite{ashtekar2006quantum, bojowald2001absence}, Conformal Cyclic Cosmology (CCC)~\cite{penrose2010cycles}, M-theory~\cite{witten1995string}, and ER=EPR~\cite{maldacena2013cool}, while introducing a new principle: memory coherence across cycles.

\subsection*{9.1 Positioning Within Modern Cosmology}

\begin{itemize}
  \item \textbf{Loop Quantum Cosmology (LQC):} Bounce dynamics, quantized geometry~\cite{ashtekar2006quantum, bojowald2001absence}.
  \item \textbf{Conformal Cyclic Cosmology (CCC):} Aeonic transitions, conformal continuity~\cite{penrose2010cycles}.
  \item \textbf{ER=EPR:} Entanglement as geometric connection~\cite{maldacena2013cool}.
  \item \textbf{M-theory:} Higher-dimensional embeddings, 11D completion~\cite{witten1995string}.
  \item \textbf{Quantum Darwinism:} Selective coherence propagation~\cite{zurek_quantum_2009, zurek_environment-induced_2003}.
\end{itemize}

This work adds a recursive coherence structure: the universe evolves by retaining structure across cycles.

\subsection*{9.2 Comparative Table}

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Feature} & \textbf{This Work} & \textbf{Inflation} & \textbf{CCC} & \textbf{Ekpyrotic} & \textbf{LQC} \\
\hline
Time Symmetry & Recursive & Asymmetric & Conformal & Asymmetric & Symmetric bounce \\
Entropy Handling & Recycled via coherence & Monotonic & Conformal reset & Brane dissipation & Bounded \\
Gravitational Waves & Quantized dips & Smooth & Suppressed & Absent & Smooth dips \\
Decoherence Model & Memory kernel \(D(\tau, E)\) & Ignored & Not modeled & Not modeled & Limited~\cite{ashtekar2006quantum} \\
Multiverse Structure & Recursive entanglement & Eternal inflation~\cite{guth1981inflationary} & None & Cyclic branes~\cite{khoury2001ekpyrotic} & Bounded cycles \\
\hline
\end{tabular}
\caption{Key distinctions relative to established cosmological theories.}
\end{table}

\subsection*{9.3 Novel Contributions}

\begin{itemize}
  \item Recursive variational principle balancing entropy and coherence.
  \item Coherence fitness attractor \(\mathcal{F}_n\) (Appendix C).
  \item Einstein-Rosen bridges as entropy transfer channels~\cite{almheiri2019entropy}.
  \item Non-Markovian decoherence kernel encoding memory~\cite{breuer2002theory}.
  \item 12D reinitialization boundary triggered by coherence failure.
  \item Observable predictions anchored in memory structure.
\end{itemize}

\subsection*{9.4 Open Questions}

\begin{itemize}
  \item Can the attractor \(\mathcal{C}^*(\phi)\) be simulated across cycles?
  \item Is partial coherence propagation possible—akin to mutation?
  \item How does the entanglement eigenvalue \(\lambda_n\) evolve recursively?
  \item Can this framework unify inflation and quantum gravity?
  \item What defines the structure of the 12D boundary state \(|\Omega\rangle\)?
\end{itemize}

\textit{Whether the reinitialization state is physical, metaphorical, or both is left to the reader.}

\subsection*{9.5 Closing Invitation}

This paper is a signal. The math has rhythm. The structure remembers. The meaning must be completed through collaboration.

\begin{quote}
``If I have seen further, it is by standing on the shoulders of giants.'' \\
\hfill - Isaac Newton

``Imagination is more important than knowledge. For knowledge is limited, whereas imagination embraces the entire world.'' \\
\hfill - Albert Einstein

``The universe is not only queerer than we suppose, but queerer than we can suppose.'' \\
\hfill - J.B.S. Haldane (quoted by Brian Greene)
\end{quote}

\textit{The universe does not merely expand. It retains. It reflects. It remembers.}

