\section*{13. Formal Structure of the Recursive Action}

\subsection*{13.1 Recursive Action Definition}

We define the total action of the multiverse as a sum over cosmological cycles:
\[
\mathcal{A}_{\text{total}} = \sum_n \mathcal{A}_n
\]

Each cycle \( n \) carries a recursive action \( \mathcal{A}_n \), evaluated over boundary configuration space \( \phi = (a, \varphi, E) \), where:
\begin{itemize}
    \item \( a \): scale factor (geometry)
    \item \( \varphi \): scalar field amplitudes (matter)
    \item \( E \): entanglement eigenvalue (quantum sector)
\end{itemize}

The action for a single cycle is:
\[
\mathcal{A}_n = \int D\phi D\phi' \left[ \Psi_n^*(\phi) K(\phi, \phi') \Psi_{n-1}(\phi') + i S_{\text{ERB}}(\phi, \phi') \right] - D[\rho_n(\phi)] + G[g_{\mu\nu}(\phi)] + O_n(\phi)
\]

\begin{itemize}
  \item \( K(\phi, \phi') \): Transition kernel from LQC; filters coherent field configurations~\cite{ashtekar2006quantum, rovelli2004quantum}.
  \item \( S_{\text{ERB}}(\phi, \phi') = \frac{A(\phi, \phi')}{4G\hbar} + i \lambda_E I(\phi, \phi') \): Action of the Einstein-Rosen bridge coupling area and mutual information~\cite{maldacena2013cool}.
  \item \( D[\rho_n] = -\text{tr}(\rho_n \ln \rho_n) + \int D(\tau, E) [O, [O, \rho(t - \tau)]] d\tau \): Decoherence functional representing entropy and recursive memory effects~\cite{breuer2002theory, grigolini1999coherence}.
  \item \( G[g_{\mu\nu}] \): Gravitational constraint term ensuring topological/geometric consistency~\cite{rovelli2004quantum}.
  \item \( O_n(\phi) \): Observer projection constraint at decoherence boundary~\cite{zurek2009quantum}.
\end{itemize}

\subsection*{13.2 Symmetries and Constraints}

This action respects a discrete time-reversal symmetry:
\[
\Psi_n(\phi) \leftrightarrow \Psi_{-n}(\phi)
\]
In absence of decoherence, recursion is symmetric in time. Coherence and memory transmission enforce a directional arrow.

All transitions obey a minimal area constraint:
\[
A(\phi, \phi') \geq \ell_{\text{Pl}}^2
\]
~\cite{bojowald2001absence, rovelli1995discreteness}

Recursive evolution is subject to:
\[
\delta \mathcal{A}_n + \lambda_C \delta(\Delta S_{\text{fwd}} - \Delta S_{\text{mem}}) = 0
\]
Here, \( \lambda_C \) serves as the informational coupling constant that balances forward entropy production with backward memory fidelity—an analog to Newton's constant for recursive coherence~\cite{gellmann1994complex}.

\subsection*{13.3 Toward Functional Dynamics}

In future work, the Euler-Lagrange equations governing recursive field evolution will be derived from variation over \( \Psi_n(\phi) \). These equations will describe:
\begin{itemize}
  \item Propagation of coherence across cycles
  \item Conditions for coherence collapse
  \item Emergence of fixed-point attractors~\cite{hartle1983wave, gellmann1994complex}
\end{itemize}

This is not a classical action of point particles in spacetime. It is a variational principle over field histories and quantum memory.

\subsection*{13.4 Interpretation: Action as Memory}

This action does not simply describe what happens. It encodes what is kept.

It is not the motion of matter that defines the universe---it is the selection of memory through coherence. At each boundary, the system interferes with itself---not only to move forward, but to refine its path.

It does not just evolve.\\
It remembers.

