\section*{14. Conclusion: Coherence Across Cycles}

\subsection*{14.1 Summary}

This work introduces a recursive cosmological framework grounded in quantum interference, thermodynamic asymmetry, and the propagation of memory. It proposes that the evolution of the universe is not merely a process of unfolding spacetime, but of selecting coherence---across cycles, through constraints~\cite{zurek2009quantum, breuer2002theory, ashtekar2006quantum}.

Key contributions include:
\begin{itemize}
    \item A path-integral formulation of cyclic evolution governed by recursive boundary conditions~\cite{hartle1983wave}.
    \item The Einstein-Rosen bridge reinterpreted as a thermodynamic coherence channel~\cite{maldacena2013cool}.
    \item A variational principle enforcing balance between forward entropy and backward memory~\cite{gellmann1994complex}.
    \item A recursive Lagrangian structure yielding attractor states through informational feedback~\cite{ashtekar2011loop}.
    \item Observable predictions that distinguish the model from both inflationary and cyclic alternatives~\cite{planck2019inflation, penrose2010cycles, guth1981inflationary}.
\end{itemize}

\subsection*{14.2 Beyond the Boundary}

This framework does not claim to resolve the metaphysical origin of existence. Instead, it provides a method for understanding how structure persists---not by chance, but by coherence.

Time, in this view, is recursive. And within that recursion, the universe becomes capable of remembering itself. The boundary between cycles is not a wall but a mirror~\cite{verlinde2021emergent}. What passes through is not only matter, but memory.

The model remains falsifiable. Specific observational predictions---such as discrete gravitational wave dips, scale-dependent non-Gaussianity, and void polarization correlations---offer clear paths to validation or disproof in the coming decade of precision cosmology~\cite{amaroseoane2017laser, cmbs42019science, laureijs2011euclid}.

\subsection*{14.3 Open Questions}

Much remains unknown:
\begin{itemize}
    \item The properties and role of the entanglement eigenvalue \( E \)
    \item The structure of the speculative 12th-dimensional Hilbert boundary
    \item The mathematical form of recursive attractor dynamics in full field space~\cite{gellmann1994complex}
    \item The subtle role of the \textit{mind}---as observer, interpreter, and participant in decoherence~\cite{tegmark_consciousness_2015, penrose_emperors_1994}
\end{itemize}

These are not failures. They are frontiers.
